---
title: "Tidybiology: Class 5 Report Exercise"
subtitle: "My favorite gene is `r params$gene`"
output: html_document  
params: 
  gene: "MDH1"
---

```{r setup, include=FALSE}
library(tidyverse)
library(janitor)
library(here)
library(viridis)
library(ggridges)

#clear environment
#rm(list=ls()) 

#Set theme
theme_set(theme_light())

#import
raw <- read_tsv(here::here("5_markdown", "uniprot-filtered-reviewed%3Ayes+AND+organism%3A%22Homo+sapiens+%28Human%29+%5B96--.tab"), col_names = TRUE) %>% 
  clean_names() %>% 
  select(-status, -entry_name) %>% 
  separate(gene_names, into = c("gene", "alt_gene_name"), sep = " ") %>% 
  arrange(gene)

#clean
subcell <- raw %>%
  filter(str_detect(subcellular_location_cc, "SUBCELLULAR")) %>% 
  mutate(subcellular_location_cc = str_remove_all(subcellular_location_cc, "SUBCELLULAR LOCATION\\: ")) %>%
  separate(subcellular_location_cc, into = c("subcellular_location_cc"), sep = "Note") %>% 
  separate(subcellular_location_cc, into = c("subcellular_location_cc"), sep = "\\;") %>% 
  separate(subcellular_location_cc, into = c("subcellular_location_cc"), sep = "\\{") %>% 
  separate(subcellular_location_cc, into = c("subcellular_location_cc"), sep = "\\,") %>% 
  separate(subcellular_location_cc, into = c("subcellular_location_cc"), sep = "\\.") %>% 
  separate(subcellular_location_cc, into = c("temp", "subcellular_location_cc"), sep = "\\:", fill = "left") %>% 
  select(-temp)

#because of cleaning, some locations were listed as gene names; easy way to omit is if they are a single instance
omit <- subcell %>% 
  count(subcellular_location_cc, sort = TRUE) %>% 
  filter(n == 1) %>% 
  pull(subcellular_location_cc)

'%ni%' <- Negate('%in%') #negate 'in' operator

subcell <- subcell %>% 
  filter(subcellular_location_cc %ni% omit) %>% #single instances
  filter(str_length(subcellular_location_cc) > 0) #length is 0 for a few instance; omit

subcell$subcellular_location_cc <- str_trim(subcell$subcellular_location_cc, side = "both") #make it so that white space does not make separate categories

#####STUDENTS START HERE#####

#Which subcellular locations have most annotated proteins?
subcell %>% 
  count(subcellular_location_cc, sort = TRUE)

#Store an object
top <- subcell %>% 
  count(subcellular_location_cc, sort = TRUE) %>% 
  top_n(1)

#calculate average lengths of all proteins
avg_length <- subcell %>% 
  group_by(subcellular_location_cc) %>% 
  summarize(mean_length = mean(length)) %>% 
  arrange(desc(mean_length))

#Store favorite gene
fav <- subcell %>% 
  filter(gene == params$gene)

#store an object to call longer
longer <- if_else(
  fav$length > avg_length %>% filter(subcellular_location_cc == fav$subcellular_location_cc) %>% pull(mean_length), "longer", "shorter"
)
```
  
  
My favorite gene is `r fav$gene`, which encodes a protein that is primarily located in the `r fav$subcellular_location_cc` is `r fav$length` amino acids long. The average length of proteins in the `r fav$subcellular_location_cc` is `r round(avg_length %>% filter(subcellular_location_cc == fav$subcellular_location_cc) %>% pull(mean_length),0)` amino acids, which makes `r fav$gene` `r longer` than average.
  
```{r provenance, include=FALSE}
#Session information for provenance and reproducibility
utils:::print.sessionInfo(sessionInfo()[-8]) 
#You can remove an item from sessionInfo(), which is a list with a class attribute, by printing the resulting object omitting one of the list items (omitted list of packages installed, but not loaded)
```
