---
title: "02_workshop"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1

You will be using the `genome` dataset to answer all the following questions

Add your answers (text and/or code chunks as required) in the space after each question

## Please answer the following questions

How would you download and read the Dplyr vignette in the Tidybio package?

```{r}

```

How would you load the `genome` datset from the Tidybio package?

```{r}

```

What type of object is `genome`?

```{r}

```

Convert `genome` to a tbl

```{r}

```


How many observations and variables does `genome` contain?

```{r}

```

What are the top five chromosomes in terms of the number of protein coding genes (look at the `protein_codinggenes` variable)? Based on your output, would you say that chromosome length is proportional to the number of protein coding genes?

```{r}

```

Using the `protein_codinggenes` and `pseudo_genes` variables, create a new variable called `total_genes` that contains the sum of protein coding genes and pseudogenes

```{r}

```

Similarly, create a new variable called `total_rna` that contains the sum of all the different rna types i.e. the summ of all variables whose name ends in `_rna`

```{r}

```

Make a table that contains the mean of `total_genes` and `total_rna`

```{r}

```

How many chromosomes have a higher than average total number of genes? Type out (or copy/paste) the names of these chromosomes

```{r}

```

## Part 2

You will be using the `uniprot_base` dataset to answer all the following questions

## Please answer the following questions

How would you load the `uniprot_base` datset from the Tidybio package?

```{r}

```

What type of object is `uniprot_base`?

```{r}

```

Convert `uniprot_base` to a tbl

```{r}

```

How many observations and variables does `uniprot_base` contain?

```{r}

```

What are the shortest and longest proteins? What are the shortest and longest genes? Are they the same?

```{r}

```


What is the difference in the lengths of the shortest and longest proteins?

```{r}

```

How many proteins of above-average length are in `uniprot_base`?

```{r}

```









