---
title: "Tidybiology Final Project"
author: "YOUR NAME HERE"
date: "`r format(Sys.time(), '%B %d, %Y')`" #put in quotes to make YAML recognize it
---
## Load libraries, such as tidyverse and tidybiology
```{r setup}

#uncomment (or select) to clear environment; good for provenance
rm(list=ls()) 
```

## Import data
DESCRIBE YOUR DATA
```{r import}
#IMPORT YOUR DATA
#MAKE SURE TO GET THE CODE CHUNK OPTIONS SET
```

## Take a `glimpse` into the data in the dataframe
DESCRIBE WHAT YOU SEE
```{r glimpse}
#DEPENDING ON DATA, NOT TOTALLY NECESSARY
```
## EDA
EXPLORATORY DATA ANALYSIS
```{r}
#COMMENCE ROLLING
#MAKE COMMENTS
#WRITE WHAT YOU SEE
```

## EDA Graph  
BEGIN TO EXPLORE HOW YOU MIGHT PLOT IT. OK TO SHOW YOUR FAILURES, OR HALF-FINISHED PRODUCTS HERE.
```{r}
#FIRST I FAILED. 
```
  
## Final Graph
SET UP THE BIG FINALE
```{r final_graph}
#MAKE IT SHINE
```

## Conclusions
WHAT WOULD YOU CONLCUDE FROM YOUR NEW OBSERVATION?
  
## Prioritized follow-up studies
WHAT WOULD YOU DO TO FOLLOW-UP?

## Acknolwedgements
SHOUT OUTS FOR PACKAGES, INSPIRATION, CLASSMATES, OR EVEN Y0UR INSTRUCTORS! ;]

#Session information for provenance and reproducibility
```{r}
#PROBABLY BEST TO LEAVE THIS HERE, BUT NO NEED TO PRINT IN THE FINAL REPORT.
utils:::print.sessionInfo(sessionInfo()[-8]) 
#You can remove an item from sessionInfo(), which is a list with a class attribute, by printing the resulting object omitting one of the list items (omitted list of packages installed, but not loaded)
```

