---
title: "Class 4- stringr workshop"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(stringr)
library(tidybiology)
library(mitocarta)
library(proteins)


```

load the proteins and mitocarta set. Join them like we did in class
```{r}

```

Now you are going to search for a new motif within the dataset. In this case, try searching for either the AcylPhosphatase 1 signature ([LIV]-x-G-x-V-Q-[GH]-V-x-[FM]-R) or the Tyrosine kinase phosphorylation site 1 motif ([RK]-x(2)-[DE]-x(3)-Y)

```{r}

```


Now add a column to the proteins dataframe using your code. Decide whether you would prefer your output to be numeric or logical. You can even try both

```{r}

```


If you have a mitochondrial (or other) protein motif of interest, feel free to explore!

A list of protein domains and motifs, along with their signatures can be found here: 
https://prosite.expasy.org/cgi-bin/prosite/prosite-list.pl 
```{r}

```

