---
title: "04_student"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Importing data into R 

The `readr` package (found in the `tidyverse` collection) contains a number of useful functions of the form  
`read_*` to import data. For example, if you have a `.csv` file, you would use the `read_csv` function

Download a file from uniprot.org

After selecting some columns of interest, click the Download button and download as a compressed Text file

Rename the file to something simple (yet informative!), like `uniprot` and make sure the extension is `.tsv`

## Importing data into R 

To import into RStudio, run the following

```{r import tsv, warning = FALSE}
uniprot <- read_tsv("uniprot.tsv")
```

## Importing data into R 

You can also use the `readr` package to import data from a URL

For example, to load a dataset from the (very useful) Tidy Tuesday series, run the following

```{r import pizza, echo = TRUE, warning = FALSE, message = FALSE, eval = TRUE}
pizza <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-01/pizza_jared.csv")
```

This data set contains ratings of various pizzerias in Manhattan

## Combining datasets 

There are many times when you have two or more overlapping datasets that you would like to combine

The `dplyr` package has a number of `*_join` functions for this purpose

## `left_join` 

Returns all rows from a, and all columns from a and b

Rows in a with no match in b will have NA values in the new columns

If there are multiple matches between a and b, all combinations of the matches are returned

## `left_join` example 

First, load the two datasets needed for this example - `proteins` and `mitocarta`

```{r load datasets, echo = TRUE, warning = FALSE}
library(proteins)
library(mitocarta)
```

Take a look at the variables in each dataset

`gene_name` in `proteins` and `symbol` in `mitocarta` contain the gene IDs for each gene. They can therefore serve as a common variable

Let's join on this

## `left_join` example 

`left_join` `proteins` with `mitocarta` and assign the output to a new object called `pm_left`

```{r left_join example, echo = TRUE, warning = FALSE}
pm_left <- left_join(proteins, mitocarta, by = c("gene_name" = "symbol"))
pm_left 
```

Now you have one dataset with additional useful information

## `right_join` 

Returns all rows from b, and all columns from a and b

Rows in b with no match in a will have NA values in the new columns

If there are multiple matches between a and b, all combinations of the matches are returned

## `right_join` example 

`right_join` `proteins` with `mitocarta` and assign the output to a new object called `pm_right`

```{r right_join example, echo = TRUE, warning = FALSE}
pm_right <- right_join(proteins, mitocarta, by = c("gene_name" = "symbol"))
pm_right 
```

## `inner_join` 

Returns all rows from a where there are matching values in b, and all columns from a and b

If there are multiple matches between a and b, all combination of the matches are returned

## `inner_join` example 

`inner_join` `proteins` with `mitocarta` and assign the output to a new object called `pm_inner`

```{r inner_join example, echo = TRUE, warning = FALSE}
pm_inner <- inner_join(proteins, mitocarta, by = c("gene_name" = "symbol"))
pm_inner 
```

Why might this type of join be useful?

## `full_join` 

Returns all rows and all columns from both a and b

Where there are no matching values, returns NA for the one missing

## `full_join` example 

`full_join` `proteins` with `mitocarta` and assign the output to a new object called `pm_full`

```{r full_join example, echo = TRUE, warning = FALSE}
pm_full <- full_join(proteins, mitocarta, by = c("gene_name" = "symbol"))
pm_full 
```



